---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Choosing colors for your chart

<!-- badges: start -->
<!-- badges: end -->

Some R code I used to generate two charts of median age and [Tree City USA](https://www.arborday.org/programs/treecityusa/) for the Carteret County, NC.
These figures were used in my [talk](https://slides.com/trang1618/ims-viz/) at the Institute of Marine Sciences at UNC Chapel Hill.

Additional scratch R scripts can be ignored.

## Load in packages

Shout out to Kyle Walker for his work on the **{tidycensus}** R package.
This package makes it extremely easy to download census data (in just a few lines like below).
Kyle also just preprinted his book [_Analyzing US Census Data: Methods, Maps, and Models in R_](https://walker-data.com/census-r/).
See an overview of the book content [here](https://twitter.com/kyle_e_walker/status/1432357038772957189).

```{r message=FALSE, warning=FALSE}
library(tidycensus)
library(tidyverse)
```

If you would like to choose a different variable (of over 20,000 variables), you can use `load_variables()` and browse the resulting dataframe.

```{r}
# v17 <- load_variables(2017, "acs5", cache = TRUE)
# View(v17)
```

## Continuous example: Age

We use the sequential color palette _viridis_ for this:

```{r carteret-age}
carteret_age_df <- get_acs(
  state = "NC", county = "Carteret", geography = "tract",
  variables = "B01002_001", geometry = TRUE
) 

carteret_age <- carteret_age_df %>%
  ggplot(aes(fill = estimate)) +
  geom_sf(color = NA) +
  coord_sf(crs = 26911) +
  scale_fill_viridis_c(
    option = "D",
    direction = -1,
    trans = "reverse"
  ) +
  labs(
    title = "Median age in Carteret County, North Carolina",
    subtitle = "2015-2019 ACS for Census Tracts",
    fill = NULL
  ) +
  theme_void() +
  theme(legend.key.height = unit(30, "pt"))

carteret_age
```

## Categorical example: Tree City USA

The census actually doesn't have this data.
I had to browse the [Tree City USA participants map](https://www.ncforestservice.gov/Urban/tcusa_NC_Participants.asp) to see which cities/census tract that participated.

```{r carteret-tree}
carteret_tree_df <- carteret_age_df %>%
  mutate(tree = if_else(grepl("9710.02|9703.04", NAME), "Tree City USA", "Nope"))

carteret_tree <- carteret_tree_df %>%
  ggplot(aes(fill = fct_rev(tree))) +
  geom_sf(color = NA) +
  coord_sf(crs = 26911) +
  scale_fill_manual(values = c("#778868", "#f6edbd")) +
  labs(
    title = "Tree Cities in Carteret County, North Carolina",
    subtitle = "2015-2019 ACS for Census Tracts",
    fill = NULL
  ) +
  theme_void()

carteret_tree
```

## Save charts

These last two lines are not run in this README, but I wanted to point out that specifying height and width would allow you to control the relative size of text vs. graph.

```{r save-charts}
# ggsave("figs/carteret_age.png", carteret_age, height = 4, width = 4.5)
# ggsave("figs/carteret-tree.png", carteret_tree, height = 4, width = 5)
```
